Equalizer\\Programming and User Guide

### 2013-01-22 Developer/ProgrammingGuide/paper.tex
Update autoconfig section

\subsubsection{GPU Service Discovery}

Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled during the
initialization of the application. The auto-configuration uses the same format
as the static configuration files. It can be used to create template
configurations by running a test application or the eqServer binary and using
the generated configuration, saved as \textless session\textgreater.auto.eqc.

Please note that both GPU-SD and the GPU-SD ZeroConf module are optional and
requires GPU-SD and a ZeroConf implementation when Equalizer is compiled.

GPU-SD contains three components: a core library, local and remote discovery
modules as well as a ZeroConf daemon. The daemon uses the appropriate module to
query local GPUs and announces them using ZeroConf on the local network.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
Equalizer uses the library with the dns\_sd discovery module to gather the
information announced by all daemons on the local network, as well as the local
discovery modules for standalone auto-configuration. The default configuration
is using all locally discovered GPUs.

\fig{fGPUSD} shows how GPU-SD is used to discover four remote GPUs on two
different machines to use them to render a single view on a laptop.

\subsubsection{\label{sGPUSDUsage}Usage}

On each node contributing to the configuration, install and start the GPU-SD
daemon. If multiple, disjoint configurations are used on the same network,
provide the session name as a parameter when starting the daemon. Verify that
all GPUs are visible on the application host using the \textsf{gpu\_sd\_list}
tool. When starting the application, use the command-line parameter
\textsf{--eq-config}:

\begin{itemize}
\item The default value is \textsf{local} which uses all local GPUs queried
  using the cgl, glx or wgl GPU-SD module.
\item \textsf{--eq-config sessionname} uses the dns\_sd ZeroConf module of
  GPU-SD to query all GPUs in the subnet for the given session. The default
  session of the GPU-SD daemon is \textsf{default}. The hostname provided by
  ZeroConf is used to connect the nodes.
\item \textsf{--eq-config filename.eqc} loads a static configuration from the
  given ASCII file. The following sections describe how to write configuration
  files.
\end{itemize}

@@

\begin{description}
\item [2D] A dynamically load-balanced sort-first configuration using all GPUs
  as sources.
\item [Simple] A no-scalability configuration using only the display GPU for
  rendering.
\item [Static DB] A static sort-last configuration distributing the range evenly
  across all GPUs.
\item [Dynamic DB] A dynamically load-balanced sort-last configuration using all
  GPUs as sources.
\end{description}

The ZeroConf hostname is used to configure the node, that is, the launch command
has to be able to resolve this hostname for starting the render client
processes. Given a properly set up ZeroConf installation, the default launch
command \textsf{ssh} should be able to start the render client correctly.

@@

\begin{description}
\item[seq] The core namespace for Sequel, the simple interface to the Equalizer
  client library.
\item[eq] The core namespace for the Equalizer client library. The classes and
  their relationship in this namespace closely model the configuration file
  format. The classes in the \textsf{eq} namespace are the main subject of this
  Programming Guide.\fig{fClientUml} provides an overview map of the most
  important classes in the Equalizer namespace, grouped by functionality.
\item[eq::util] The eq::util namespace provides common utility classes, which
  often simplify the usage of OpenGL functions. Most classes in this namespace
  are used by the Equalizer client library, but are usable independently
  from Equalizer for application development.
\item[eq::admin] The eq::admin namespace implements an administrative API to
  change the configurations of a running server. This admin API is not yet
  finalized and will very likely change in the future.
\item[eq::fabric] The eq::fabric namespace is the common data management and
  transport layer between the client library, the server and the administrative
  API. Most Equalizer client classes inherit from base classes in this namespace
  and have all their data access methods in these base classes.
\item[co] Collage is the network library used by Equalizer. It provides basic
  functionality for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{LocalNode}, \textsf{Object} and
  \textsf{Serializable}. Please refer to \sref{sNetwork} for an introduction
  into the network layer, and to \sref{sNetObject} for distributed objects.
\item[lunchbox] The lunchbox library provides C++ classes to abstract the
  underlying operating system and to implement common helper functionality for
  multi-threaded applications. Examples are \textsf{lunchbox::Clock} providing a
  high-resolution timer, or \textsf{lunchbox::MTQueue} providing a thread-safe,
  blocking FIFO. Classes in this namespace are fully documented in the API
  documentation on the Equalizer website, and are not subject of this
  Programming Guide.
\item[hwloc, boost, vmmlib, gpu-sd] External libraries providing manual and
  automatic thread affinity, serialization and the foundation for RSP multicast,
  vector and matrix mathematics as well as local and remote GPU discovery,
  respectively. The hwloc and gpu-sd libraries are used only internally and not
  exposed through the API.
\item[eq::server] The server namespace, implementing the functionality of the
  Equalizer server, which is an internal namespace not to be used by application
  developers. The \textsf{eq::admin} namespace enables run-time administration
  of Equalizer servers. The server does not yet expose a stable
  API.
\end{description}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/namespaces}
  {\caption{\label{fNamespaces}Namespaces}}
\end{wrapfigure}
The Equalizer examples are implemented in their own namespaces, e.g.,
\textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing from the
\textsf{eq} namespace, with the occasional usage of functionality from the
\textsf{eq::util}, \textsf{eq::fabric}, \textsf{co} and \textsf{lunchbox}
namespace. \fig{fNamespaces} shows the namespaces and their layering.

@@

The application's main loop starts by connecting the application to an Equalizer
server. If no server is specified, \textsf{Client::connectServer} tries first to
connect to a server on the local machine using the default port. If that fails,
it will create a server running within the application process using
auto-configuration as described in \sref{sGPUSDUsage}. The command line
parameter \textsf{--eq-config} can be used to specify a gpu-sd session or
configuration file, and \textsf{--eq-server} to explicitly specify a server
address.

@@

\begin{description}
\item[--eq-help] shows all available library command line options and their
  usage.
\item[--eq-client] is used to start render client processes. Starts a resident
  render client when used without additional arguments, as described in
  \sref{sResident}. The server appends an undocumented argument to this option
  which is used by the \textsf{eq::Client} to connect to the server and identify
  itself.
\item[--eq-layout \textless layoutName\textgreater] activates all layouts of the
  given name on all respective canvases during initialization. This option may
  be used multiple times.
\item[--eq-modelunit \textless unitValue\textgreater] is used for scaling the
  rendered models in all views. The model unit defines the size of the model wrt
  the virtual room unit which is always in meter. The default unit is 1 (1 meter
  or EQ\_M).
\item[--eq-logfile \textless filename\textgreater] redirects all Equalizer
  output to the given file.
\item[--eq-server \textless connection\textgreater] provides a connection
  description when using a separate server process.
\item[--eq-config \textless sessionName\textbar filename.eqc\textgreater] is
  used to configure an application-local server, as described in
  \sref{sGPUSDUsage}.
\item[--eq-config-flags \textless multiprocess\textbar multiprocess\_db\textbar
  ethernet\textbar infiniband\textbar 2D\_horizontal\textbar
  2D\_vertical\textbar 2D\_tiles\textgreater] is used as input for the
  autoconfiguration to tweak the configuration generation.
\item[--eq-config-prefixes \textless CIDR-prefixes\textgreater] is used as input
  for the autoconfiguration to filter network interfaces.
\item[--eq-render-client \textless filename\textgreater] provides an alternate
  executable name for starting render clients. This option is useful if the
  client machines use a different directory layout from the application machine.
\item[--eq-listen \textless connection\textgreater] configures the
  local node to listen on the given connections. The formation of the connection
  argument is documented in
  \textsf{co::Con\-nec\-tionDescription::fromString()}. This option may be used
  multiple times.

>>>>

\subsubsection{Hardware Service Discovery}

Equalizer uses the hwsd library for discovering local and remote GPUs and
network interfaces. Based on this information, a typical configuration is
dynamically compiled during the initialization of the application. The
auto-configuration uses the same format as the static configuration files. It
can be used to create template configurations by running a test application or
the eqServer binary and using the generated configuration, saved as
\textless session\textgreater.auto.eqc.

Please note that both hwsd and the hwsd ZeroConf module are optional and
requires hwsd and a ZeroConf implementation when Equalizer is compiled.

hwsd contains three components: a core library, local and remote discovery
modules as well as a ZeroConf daemon. The daemon uses the appropriate module to
query local GPUs and network interfaces and announces them using ZeroConf on the
local network.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf} % TODO: update image (hwsd)
  {\caption{\label{fHWSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
Equalizer uses the library with the dns\_sd discovery module to gather the
information announced by all daemons on the local network, as well as the local
discovery modules for standalone auto-configuration. The default configuration
is using all locally discovered GPUs and network interfaces.

\fig{fHWSD} shows how hwsd is used to discover four remote GPUs on two
different machines to use them to render a single view on a laptop.

\subsubsection{\label{sHWSDUsage}Usage}

On each node contributing to the configuration, install and start the hwsd
daemon. If multiple, disjoint configurations are used on the same network,
provide the session name as a parameter when starting the daemon. Verify that
all GPUs and network interfaces are visible on the application host using
the \textsf{hw\_sd\_list} tool. When starting the application, use the
command-line parameter \textsf{--eq-config}:

\begin{itemize}
\item The default value is \textsf{local} which uses all local GPUs and network
  interfaces queried using the cgl, glx, or wgl GPU module and the sys network
  module.
\item \textsf{--eq-config sessionname} uses the dns\_sd ZeroConf module of
  hwsd to query all GPUs and network interfaces in the subnet for the given
  session. The default session of the hwsd daemon is \textsf{default}. The
  found network interfaces are used to connect the nodes.
\item \textsf{--eq-config filename.eqc} loads a static configuration from the
  given ASCII file. The following sections describe how to write configuration
  files.
\end{itemize}

@@

\begin{description}
\item [2D] A dynamically load-balanced sort-first configuration using all GPUs
  as sources.
\item [Simple] A no-scalability configuration using only the display GPU for
  rendering.
\item [Static DB] A static sort-last configuration distributing the range evenly
  across all GPUs.
\item [Dynamic DB] A dynamically load-balanced sort-last configuration using all
  GPUs as sources.
\item [DB Direct Send] A direct send sort-last configuration using all GPUs as
  sources.
\item [DB 2D] A direct send sort-last configuration using all nodes together
  with a dynamically load-balanced sort-first configuration using local GPUs.
\end{description}

All suitable network interfaces are used to configure the node, that is, the
launch command has to be able to resolve one hostname for starting the render
client processes. Suitable interfaces have to be up and match optional given
values which can be specified by the following command-line parameters:

\begin{description}
\item[--eq-config-flags \textless ethernet\textbar infiniband\textgreater]
  Limit the selection of network interfaces to one of the those types.
\item[--eq-config-prefixes \textless CIDR-prefixes\textgreater] Limit the
  selection of network interfaces that match the given prefix.
\end{description}

@@

\begin{description}
\item[seq] The core namespace for Sequel, the simple interface to the Equalizer
  client library.
\item[eq] The core namespace for the Equalizer client library. The classes and
  their relationship in this namespace closely model the configuration file
  format. The classes in the \textsf{eq} namespace are the main subject of this
  Programming Guide.\fig{fClientUml} provides an overview map of the most
  important classes in the Equalizer namespace, grouped by functionality.
\item[eq::util] The eq::util namespace provides common utility classes, which
  often simplify the usage of OpenGL functions. Most classes in this namespace
  are used by the Equalizer client library, but are usable independently
  from Equalizer for application development.
\item[eq::admin] The eq::admin namespace implements an administrative API to
  change the configurations of a running server. This admin API is not yet
  finalized and will very likely change in the future.
\item[eq::fabric] The eq::fabric namespace is the common data management and
  transport layer between the client library, the server and the administrative
  API. Most Equalizer client classes inherit from base classes in this namespace
  and have all their data access methods in these base classes.
\item[co] Collage is the network library used by Equalizer. It provides basic
  functionality for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{LocalNode}, \textsf{Object} and
  \textsf{Serializable}. Please refer to \sref{sNetwork} for an introduction
  into the network layer, and to \sref{sNetObject} for distributed objects.
\item[lunchbox] The lunchbox library provides C++ classes to abstract the
  underlying operating system and to implement common helper functionality for
  multi-threaded applications. Examples are \textsf{lunchbox::Clock} providing a
  high-resolution timer, or \textsf{lunchbox::MTQueue} providing a thread-safe,
  blocking FIFO. Classes in this namespace are fully documented in the API
  documentation on the Equalizer website, and are not subject of this
  Programming Guide.
\item[hwloc, boost, vmmlib, hwsd] External libraries providing manual and
  automatic thread affinity, serialization and the foundation for RSP multicast,
  vector and matrix mathematics as well as local and remote hardware (GPU,
  network interfaces) discovery, respectively. The hwloc and hwsd libraries are
  used only internally and not exposed through the API.
\item[eq::server] The server namespace, implementing the functionality of the
  Equalizer server, which is an internal namespace not to be used by application
  developers. The \textsf{eq::admin} namespace enables run-time administration
  of Equalizer servers. The server does not yet expose a stable
  API.
\end{description}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/namespaces} % TODO: update image (hwsd)
  {\caption{\label{fNamespaces}Namespaces}}
\end{wrapfigure}
The Equalizer examples are implemented in their own namespaces, e.g.,
\textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing from the
\textsf{eq} namespace, with the occasional usage of functionality from the
\textsf{eq::util}, \textsf{eq::fabric}, \textsf{co} and \textsf{lunchbox}
namespace. \fig{fNamespaces} shows the namespaces and their layering.

@@

The application's main loop starts by connecting the application to an Equalizer
server. If no server is specified, \textsf{Client::connectServer} tries first to
connect to a server on the local machine using the default port. If that fails,
it will create a server running within the application process using
auto-configuration as described in \sref{sHWSDUsage}. The command line
parameter \textsf{--eq-config} can be used to specify a hwsd session or
configuration file, and \textsf{--eq-server} to explicitly specify a server
address.

@@

\begin{description}
\item[--eq-help] shows all available library command line options and their
  usage.
\item[--eq-client] is used to start render client processes. Starts a resident
  render client when used without additional arguments, as described in
  \sref{sResident}. The server appends an undocumented argument to this option
  which is used by the \textsf{eq::Client} to connect to the server and identify
  itself.
\item[--eq-layout \textless layoutName\textgreater] activates all layouts of the
  given name on all respective canvases during initialization. This option may
  be used multiple times.
\item[--eq-modelunit \textless unitValue\textgreater] is used for scaling the
  rendered models in all views. The model unit defines the size of the model wrt
  the virtual room unit which is always in meter. The default unit is 1 (1 meter
  or EQ\_M).
\item[--eq-logfile \textless filename\textgreater] redirects all Equalizer
  output to the given file.
\item[--eq-server \textless connection\textgreater] provides a connection
  description when using a separate server process.
\item[--eq-config \textless sessionName\textbar filename.eqc\textgreater] is
  used to configure an application-local server, as described in
  \sref{sHWSDUsage}.
\item[--eq-config-flags \textless multiprocess\textbar multiprocess\_db\textbar
  ethernet\textbar infiniband\textbar 2D\_horizontal\textbar
  2D\_vertical\textbar 2D\_tiles\textgreater] is used as input for the
  autoconfiguration to tweak the configuration generation.
\item[--eq-config-prefixes \textless CIDR-prefixes\textgreater] is used as input
  for the autoconfiguration to filter network interfaces.
\item[--eq-render-client \textless filename\textgreater] provides an alternate
  executable name for starting render clients. This option is useful if the
  client machines use a different directory layout from the application machine.
\item[--eq-listen \textless connection\textgreater] configures the
  local node to listen on the given connections. The formation of the connection
  argument is documented in
  \textsf{co::Con\-nec\-tionDescription::fromString()}. This option may be used
  multiple times.
